---
title: "Activity Monitoring Data"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

<style type="text/css">
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
</style>


### Loading and preprocessing the data


```{r}
library( ggplot2)
df=read.csv("./data/repdata_data_activity/activity.csv")

str(df)

df$date=as.POSIXct(df$date)
```

### What is mean total number of steps taken per day?

```{r}
steps_daily=aggregate(df$steps~df$date,FUN = sum)
colnames(steps_daily)=c("Date","Steps")


ggplot(steps_daily,aes(x=Steps))+
  geom_histogram()+
  ggtitle("Steps Taken Daily")+
  xlab("Steps")

  


```


### Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean_old=mean(steps_daily$Steps,na.rm = T)
median_old=median(steps_daily$Steps,na.rm = T)
print(paste("Mean is:",mean_old,"&  Median is:",median_old))

```

### What is the average daily activity pattern?

```{r}

meansteps=aggregate(df$steps~df$interval,FUN = mean)
colnames(meansteps)=c("Interval","Steps")

ggplot(meansteps,aes(x=Interval,y=Steps))+
  geom_line()+
  ggtitle("Average Daily Activity")

```


### Total missing values

```{r}
missingvalues = nrow(df[is.na(df$steps),])
missingvalues

```

### Devise a strategy for filling in all of the missing values in the dataset


```{r}
df_filled <- transform(df,steps = ifelse(is.na(df$steps),
                          meansteps$Steps[match(df$interval, 
                                              meansteps$Interval)],
                                             df$steps))

df_filled_agg <- aggregate(df_filled$steps~df_filled$date, FUN=sum)
colnames(df_filled_agg)=c("Date","Steps")

hist(df_filled_agg$Steps,
     main = "Number of Steps Per Day",
     xlab = "Number of Steps")


```


### Calculate and report the mean and median total number of steps taken per day.

```{r}
mean_new=mean(df_filled_agg$Steps,na.rm = T)
median_new=median(df_filled_agg$Steps,na.rm = T)
print(paste("Mean is:",mean_new,"&  Median is:",median_new))

```

### Do these values differ from the estimates from the first part of the assignment? 

```{r}
print(paste("Difference in mean:",mean_new-mean_old))
print(paste("Difference in median:",median_new-median_old))
```

### Are there differences in activity patterns between weekdays and weekends?

```{r}


# Let's create defferciate bw weekday and weekend
Type_of_day <- function(date) {
  day_name <- weekdays(date)
  if (day_name %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return ("Weekday")
  else if (day_name %in% c('Saturday', 'Sunday'))
      return ("Weekend")
}

df_filled$day <- sapply(df_filled$date, FUN = Type_of_day)



```


### Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}

stepsbyDaytype <- aggregate(steps ~ interval + day, df_filled, mean)
ggplot(data = stepsbyDaytype, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  
  xlab("Interval") +
  ylab("Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))


```

